The Shape Matching Element Method for Meshless Animation of NURBs Models

### 2021-01-23 paper/sections/methods.tex
blending weights section done

\subsection{Choosing Deformation Centers}
\label{sec:origins}
\dave{Ty: try to follow this formula when writing 
(1) what is the purpose of this part of the method ?
(2) what are the criteria for success
(3) what is our assumed input 
(4) what is the desired output
(5) what's the algorith,
(6) add proofs/diagrams/figures that show we satisfy (2). }

\subsection{Meshless Blending Weight Computation}
\label{sec:weights}
\dave{Desireable Properties}
\dave{Construction via post normalization}
\dave{only need blending weights at quadrature points and surface samples so given sparse set of quadrature points compute blending weights using raycasting}
\dave{things to explain (1) raycasting works for overlapping geometric, seperated geometry (2) weights glue objects together}
\dave{maybe but weight figure here so reviewers immediately know that it looks pretty good}
\dave{need to mention Monte Carlo Geometry processing but can't use it because our parts can overlap and that's a tricky boundary condition}

All that remains for the deformation map is to compute blending weights. Our choice of the weighting function is guided by the simple idea that that for a given material point its weight should be highest to whichever surfaces it is closest. This idea lends itself to a simple distance-based weighting method. The first step in this process is to compute \textit{distance weights}. For a given point $\mathbf{X}$ and a set of $n$ we wish to compute a set of distance weights to each surface $\mathbf{\theta}(\mathbf{X}) = \left(\theta_1, \dots, \theta_i, \dots, \theta_n \right)$.

For the $i$-th surface, we first find the distance, $d_{\text{primary}}$, from the point $\mathbf{X}$ to the closest point, $\mathbf{p}$, on the surface. Next we wish to satisfy the condition that the distance weight is highest on the surface and it decays towards zero as it approaches other surfaces. To achieve this, we shoot a ray from $\mathbf{p}$ towards $\mathbf{X}$ and check for intersections between this ray and the other $n-1$ surfaces. If this ray hits another surface, we let $d_{\text{total}}$ be the distance from $\mathbf{p}$ to the closest ray hit $\mathbf{q}$. These two distances enable us to write a simple linear function to compute the \textit{distance weight}:
\begin{equation}
\theta_i = \max (1.0 - \frac{d_{\text{primary}}}{\min (d_{\text{total}}, D)}, 0.0)
\text{,}
\end{equation}
where $D$ is a fixed cutoff distance parameter. We see that this distance weight is the result of interpolating along this ray from $\mathbf{p}$ to $\mathbf{q}$ (or a fixed point that depends on $D$). Importantly, this gives a result where the distance weight to surface $i$ is equal to $1$ if it lies on $i$ and $0$ if it lies on another surface. More elaborate schemes may be developed for computing distances, such as using the Boundary Element Method, but we find this simple linear function to perform well for our purposes.

\begin{figure}
    \includegraphics[width=\columnwidth]{example-image-a}
    \caption{Figure showing how weight calculation works (left) ray casting plus result (right) after correcting for partition of unity}
    \label{fig:weightcompute}
\end{figure}

Now that we have a set of distances weights $\mathbf{theta}$, we seek to form a final set of \textit{blending weights} $\mathbf{w}(\mathbf{X}) =  \left( w_1, \dots, w_i, \dots, w_n \right)$ to blend the polynomials of the $n$ shape elements. In constructing these blending weights, we require that they adhere to a set of conditions. We require that (1) each weight is non-negative, (2) the sum of the weights is equal to 1 (\textit{partition of unity}), (3) and that if a point lies on a surface $i$, it will have a weight of 1 to $i$ and consequently zero to the others. To produce these blending weights we form a quadratic program that satisfies these constraints:
\begin{equation}
\begin{aligned}
\mathbf{w}(\mathbf{X}) = \min_{\mathbf{w}} \quad & \mathbf{w}^T \Theta(\mathbf{X}) \mathbf{w}    \\
\textrm{s.t.} \quad & 0 \leq w_i \leq 1                     \\
                    &   \sum_i^n w_i = 1                      \\
\end{aligned}
\end{equation}

where
\begin{equation}
     \Theta(\mathbf{X}) = \text{diag}\left( \frac{1}{\theta_1},\frac{1}{\theta_2},\dots,\frac{1}{\theta_n}\right)
\end{equation}
We see that through the use of linear constraints we satisfy conditions (1) and (2). Satisfying the third condition (3) is achieved through using the distance weights. Intuitively, we see that for some small $\theta_i$, $\frac{1}{\theta_i}$ will be very large, encouraging the quadratic program to assign a small weight for $w_i$. Conversely, we see the opposite effect for large values of $\theta_i$.

>>>>

\subsection{Choosing Deformation Centers}
\label{sec:origins}
\dave{Ty: try to follow this formula when writing 
\dave{maybe this should come after weight computation ?}
(1) what is the purpose of this part of the method ?
(2) what are the criteria for success
(3) what is our assumed input 
(4) what is the desired output
(5) what's the algorith,
(6) add proofs/diagrams/figures that show we satisfy (2). }

\subsection{Meshless Blending Weight Computation}
\label{sec:weights}
% \dave{Desireable Properties}
% \dave{Construction via post normalization}
% \dave{only need blending weights at quadrature points and surface samples so given sparse set of quadrature points compute blending weights using raycasting}
% \dave{things to explain (1) raycasting works for overlapping geometric, seperated geometry (2) weights glue objects together}
% \dave{maybe but weight figure here so reviewers immediately know that it looks pretty good}
% \dave{need to mention Monte Carlo Geometry processing but can't use it because our parts can overlap and that's a tricky boundary condition}

Our kinematic map~(\refeq{kinematics}) requires blending weights in order to homogenize our per-part representation. 
Specifying weights manually would be antithetical to our goal of providing a seamless translation from surface representation to physics-based animation.
Automatic  weight computation is a well-studied topic (see ~\citet{10.1145/2766952,BBW:2011} and ~\citet{10.1145/2010324.1964968}) but typically requires solving 
an optimization problem that itself relies on a volumetric discretization. 
This can be avoided by using stochastic methods~\cite{Sawhney:2020:MCG} but these do not yet support the boundary conditions we require for our problem.

Ideally our shape functions would both partition unity and obey the Kronecker delta property (the $i^th$ blending function is one at $i^th$ part and zero on all other parts).
These properties ensure that our simulation can properly represent translation and that the solution stored at each part is the actual deformed position.
Additionally we would like our weights to be sparse both for performance reasons (leads to sparse matrices) and for expressivity (leads to locality of deformation).

While partition of unity and sparsity are straightforward to enforce, the overlapping and intersecting parts~(\reffig{badmodels}) often encountered in non-engineering models leads
to a contradiction -- how can the blending functions be both one and zero at the points of overlap ? 
Our solution is to allow weights to include bounded discontinuities at these points and it is these discontinuous boundary conditions that are difficult to enforce 
with previous methods. 
\dave{ideally need a didactic figure showing this}

Given a set of reference space points $\st{X}$ inside our model, we want to evaluate $\bld_i\left(\refX_j\right)$, where $\bld_i$ is the weight function associated with 
the $i^{th}$ part, $\prt_i$, and $\refX_j\in\st{X}$.

For each $\refX_j$ we find the nearest point on $\prt_i$, $\clst$.
We cast a ray from $\clst$, towards $\refX_j$ and find the intersection with the nearest part, $\rhit$.

\dave{Ty: How do we deal with non-planar parts ? What if the ray cast hits the same part that the closest point is on?}

Next we compute two distances, $d_{\text{primary}}$, the distance from $\mathbf{X}_j$ to $\clst$ and $d_{\text{total}}$ the distance from  $\clst$ to $\rhit$.
Using these two distances we define a linear \emph{distance weight} using the well-known ReLU function:
\begin{equation}
\theta_i = \max (1.0 - \frac{d_{\text{primary}}}{\min (d_{\text{total}}, \alpha)}, 0.0)
\text{,}
\end{equation}
where $\alpha$ is a distance cutoff parameter. 
This distance weight~\reffig{plot_distance_weight} is the result of interpolating along this ray from $\clst$ to $\rhit$ (or a fixed point that depends on $\alpha$). 
These weights obey the Kronecker delta property for all parts but does not partition unity.

\begin{figure}[h]
    \centering
    \includegraphics[width=3.33in]{figures/plot_distance_weight.pdf}
    \caption{We plot the distance weight with respect to the primary and total ray length. Here the cutoff distance is 50.
    }
    \label{fig:plot_distance_weight}
    \vspace{-5pt}
  \end{figure}

%\begin{figure}
%    \includegraphics[width=\columnwidth]{example-image-a}
%    \caption{Figure showing how weight calculation works (left) ray casting plus result (right) after correcting for partition of unity}
%    \label{fig:weightcompute}
%\end{figure}

Next we use these distance weights to compute our final blending weights.
At each $\refX_j$ we want to compute $\mathbf{\bld}(\mathbf{X}_j) =  \left( \bld_1(\mathbf{X}_j), \dots, \bld_n(\mathbf{X}_j) \right)$, where 
$\sum_{i=1}^n\bld_i(\mathbf{X}_j) = 1$, and each $\bld_i \geq 0$, which we enforce by solving a local optimization problem:

\begin{equation}
\begin{aligned}
\mathbf{w}(\mathbf{X}_j) =  \arg\min_{\mathbf{w}} \quad & \mathbf{w}^T \Theta(\mathbf{X}_j) \mathbf{w}    \\
\textrm{st} \quad & 0 \leq w_i \leq 1 \mbox{ }\forall i                    \\
                  &   \sum_i^n w_i = 1                      \\
\end{aligned}
\end{equation}

where $\Theta(\mathbf{X}) = \text{diag}\left( \frac{1}{\theta_1},\frac{1}{\theta_2},\dots,\frac{1}{\theta_n}\right)$.

This post-process ensures that our blending weights partition unity and are non-negative~(\reffig{distance_weight_puft}). 
For any point on a part that is non-intersecting, the weights will satisfy the Kronecker delta property since the inverse distance weights imply that only
a single theta will have a non-infinite weight. 
For points that lie on more than one part, the weights associated with the intersecting parts will be equal and all others will be zero
In other words, the deformed positions of these intersecting points will be an average, which is a plausible solution and ensures continuity.
The blending weights immediately jump back to one and zero for non-intersecting points on each part. 
Finally, the distance cutoff $\alpha$ allows us to control the sparsity of the blending weights.

\begin{figure}[h]
    \centering
    \includegraphics[width=3.33in]{figures/distance_weight_puft.pdf}
    \caption{Our blending weights decay smoothly from 1.0 (yellow) to 0.0 (green) when moving away from its closest surface. Here we visualize the distribution of the distance weights (with cutoff distance 5.0) corresponding to each part.   
    }
    \label{fig:distance_weight_puft}
    \vspace{-5pt}
  \end{figure} 
  %
  %

Using raycasting to compute blending weights imbues our method with similar advantages to \citet{Sawhney:2020:MCG} (allowing us to 
handle intersecting geometry, geometry with gaps and to perform constructive solid geometry operations on the fly),
but allows us to use more general weighting functions and apply the appropriate behavior for intersecting parts.
The price we pay for these advantages is that our weights are not C0 continuous but rather Lipchitz smooth.
Perturbations in the boundaries of parts can cause bounded jumps in our distance weights. 
However these jumps occupy an infinitesimal percentage of the volume of our object are easy to avoid when integrating
physical quantities, leaving our simulation unaffected.

### 2021-01-23 paper/sections/results.tex
Added raycasting and initial patch test figure
added initial mug figure. needs work
blending weights section done

\dave{we should show the nurbs model in rhino/fusion for every example, maybe an exploded view as well}
\dave{mention that not so many standard NURBS models so we modelled lots of these ourselves .. show's power of approach that we can model and sim all these examples.}
\dave{use as many different meshes as possible for figures (i.e try not to have the same mesh show up in more than one figure if possible).}

@@

%% Results that relate to preprocessing
% weights are good
\begin{figure}
  \centering
  \includegraphics[width=3.33in]{figures/distance_weight_puft.pdf}
  \caption{Our distance weights decay smoothly from 1.0 (yellow) to 0.0 (green) when moving away from its closest surface. Here we visualize the distribution of the distance weights (with cutoff distance 5.0) corresponding to each part.   
  }
  \label{fig:distance_weight_puft}
  \vspace{-5pt}
\end{figure} 
%
%
\begin{figure}
  \centering
  \includegraphics[width=3.33in]{figures/plot_distance_weight.pdf}
  \caption{We plot the distance weight with respect to the primary and total ray length. Here the cutoff distance is 50.
  }
  \label{fig:plot_distance_weight}
  \vspace{-5pt}
\end{figure}

@@

%% Results that relate to simulation output
%correctness
\begin{figure}
  \includegraphics[width=\columnwidth]{example-image-a}
  \caption{2D examples (left) rigidly transform boundary and show that solving the static problem gives us rigid motion on the interior (right) isotropically scale boundary and show solving static problem gives us constant strain on the interior}
  \label{fig:patchtest}
\end{figure}

@@

% relaxed modeling  constraints
\begin{figure}
  \includegraphics[width=\columnwidth]{example-image-a}
  \caption{Coffee mug with overlapping handle (top) row of models with increasing overlap (middle) single weight image showing behaviour in the overlapping region, (bottom) simulation result for each one}
  \label{fig:badmodels}
\end{figure}

>>>>

\dave{we should show the nurbs model in rhino/fusion for every example, maybe an exploded view as well}

\dave{mention that not so many standard NURBS models so we modelled lots of these ourselves .. show's power of approach that we can model and sim all these examples.}

\dave{use as many different meshes as possible for figures (i.e try not to have the same mesh show up in more than one figure if possible).}

\dave{we should report which models we made ourselves}

@@

%% Results that relate to preprocessing
% weights are good
% \begin{figure}
%   \centering
%   \includegraphics[width=3.33in]{figures/distance_weight_puft.pdf}
%   \caption{Our distance weights decay smoothly from 1.0 (yellow) to 0.0 (green) when moving away from its closest surface. Here we visualize the distribution of the distance weights (with cutoff distance 5.0) corresponding to each part.   
%   }
%   \label{fig:distance_weight_puft}
%   \vspace{-5pt}
% \end{figure} 
% %
% %
% \begin{figure}
%   \centering
%   \includegraphics[width=3.33in]{figures/plot_distance_weight.pdf}
%   \caption{We plot the distance weight with respect to the primary and total ray length. Here the cutoff distance is 50.
%   }
%   \label{fig:plot_distance_weight}
%   \vspace{-5pt}
% \end{figure}

@@

%% Results that relate to simulation output
%correctness
\begin{figure}
  \includegraphics[width=\columnwidth]{figures/patch_test.pdf}
  \caption{2D examples (left) rigidly transform boundary and show that solving the static problem gives us rigid motion on the interior (right) isotropically scale boundary and show solving static problem gives us constant strain on the interior}
  \label{fig:patchtest}
\end{figure}

@@

% relaxed modeling  constraints
\begin{figure}
  \includegraphics[width=\columnwidth]{figures/mug_overlap}
  \caption{Coffee mug with overlapping handle (top) row of models with increasing overlap (middle) single weight image showing behaviour in the overlapping region, (bottom) simulation result for each one}
  \label{fig:badmodels}
\end{figure}
